%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Numerical Linear Algebra class winter semester 2023/2024
% Sheet 1                             
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{Sheet}[to be handed in until October 25, 2023, 11am.]
  \label{sheet1}
  
  Please hand in your solutions in groups of two to four. 
  The solutions for non-programming problems and the discussion of the programming problems should be handed in on paper (Mathematikon, 1st floor, letterbox 27). 
  The code for programming problems should be handed in via Moodle by one person in your group. 
  To recieve the points for the exercises on Müsli, write everyones first and last name on every hand-in. 
  Details will be discussed in the exercise groups on Friday. 
  If you currently don’t have access to Moodle, please contact your tutor Laura in the exercise group (or via Müsli/Mail). 
  If you want to get comments on your code, please also print it out and hand it in with the other solutions.
  
\begin{Problem}
  Review the following items and write down at least one of the
  definitions and one of the theorems with proof in detail.
  \begin{itemize}
  \item Definition of a projection
  \item Definition of an orthogonal projection
  \item Theorem: Orthogonal projection is uniquely determined by
    subspace

    Consider a finite-dimensional space $V$ with inner product
    $\scal(\cdot,\cdot)$ and a subspace $W\subset V$. Then, there
    exists a unique orthogonal projection
    \begin{gather*}
       P_W:V\to W.
    \end{gather*}
  \item Theorem: Best Approximation Theorem

    Let $W$ be a subspace of $\R^n$, let $\vx$ be any vector in
    $\R^n$, and let $\tilde{\vx}$ be the orthogonal projection of
    $\vx$ onto $W$. Then $\tilde{\vx}$ is the closest point in $W$ to
    $\vx$, in the sense that
    \begin{gather*}
      \norm{\vx - \tilde{\vx}} < \norm{\vx - \vw}
    \end{gather*}
    for all $\vw$ in $W$ distinct from $\vx$.
  \item Theorem: Orthogonal projection in orthonormal basis
    
    Let $B = \{\vu_1,\vu_2, ...,\vu_p\}$ be an orthonormal basis of a
    subspace $W$ of a finite-dimensional space $V$ with inner product
    $\scal(\cdot,\cdot)$. Then, the orthogonal projection $P_i$ of any
    vector $\vv\in V$ onto $\vu_i$, and the orthogonal projection
    $P_W$ of any vector $\vv\in V$ onto $W$ have the following
    expressions, respectively:
    \begin{gather*}
      P_i(\vv) = \scal(\vv, \vu_i) \vu_i,\qquad i = 1, 2, ..., p,
      \\
      P_W(\vv) = \sum_{i=1}^p \scal(\vv, \vu_i) \vu_i,
    \end{gather*}
    and
    \begin{gather*}
      \vv = P_W(\vv) + \vz,\qquad \vz \perp W.
    \end{gather*}

  \item Theorem: Parseval identity

    Suppose that $W$ is a finite-dimensional linear space with inner
    product $\scal(\cdot,\cdot)$. Let $\{\ve_i\}$, $i=1,...,n$ be an
    orthonormal basis of $W$. Then, for every $\vw\in W$ it holds
    \begin{gather*}
      \sum_{i=1}^n \abs{\scal(\vw,\ve_i)}^2
      = \norm{\vw}^2.
    \end{gather*}
  \end{itemize}
\end{Problem}

\begin{Problem}[Programming]
  Problem 1.1.4 in the Lecture Notes.
\end{Problem}

\end{Sheet}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: 
