%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Numerical Linear Algebra class 2024/25 
% Sheet 10
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{Sheet}[discussion in the first tutorials of January, 2025]
  \label{sheet10}

  This exercise sheet reviews some topics of this lecture. 
  Try first to answer withour looking at the notes.
  Your answers do not have to be handed in, 
  but the exercises will be discussed in the tutorials of the first week of January.

  \begin{Problem}
    Recapitulate the concept of an orthogonal projection and an
    oblique projection. What are use cases of both and why are they
    important for numerical linear algebra?
  \end{Problem}

  \begin{Problem}
    How can the QR factorization of a given matrix be computed?
    Discuss downsides and benefits of the different methods.
  \end{Problem}

  \begin{Problem}
    Householder reflections vs. Givens rotations: Which one is more
    cost-efficient in the general case? 
    When is using the other one advantageous?
  \end{Problem}

  \begin{Problem}
    Which of the discussed methods for solving eigenvalue problems can
    be implemented without explicitly forming a matrix?
  \end{Problem}

  \begin{Problem}
    Consider an $n\times n$ matrix that has $n$ distinct eigenvalues
    such that $\abs{\lambda_i} \neq \abs{\lambda_j}$ for $i\neq
    j$. How can the eigenvalue with the second largest absolute value
    be computed?
  \end{Problem}

  \begin{Problem}
    When does one step of the Gauss-Seidel iteration provide the
    direct solution of a linear system? Consider an upper triangular
    matrix to visualize this matter.
  \end{Problem}

  \begin{Problem}
	  What are drawbacks of the orthogonal subspace iteration?
	  When does it converge?
  \end{Problem}

  \begin{Problem}
	  Why is the implicit double shift better than the explicit?
  \end{Problem}

  \begin{Problem}
	  What is deflation? Why do we use it?
  \end{Problem}

  \begin{Problem}
	  What is polynomial filtering?
  \end{Problem}
	
  \begin{Problem}
	  Why do we use iterative solvers?
  \end{Problem}

  \begin{Problem}
	  How does the Gauss-Seidel method converge for an upper triangular matrix?
  \end{Problem}

	  \begin{Problem}
		  What is the difference between the steepest descent and the minimal residual method?
	  \end{Problem}
  
	  \begin{Problem}
		  What is the difference between the steepest descent and the conjugate gradient method?
	  \end{Problem}

	  \begin{Problem}
		  Why do we use conjugate gradient and not a straight forward implementation of a Lanczos solver?
	  \end{Problem}

	  \begin{Problem}
		  What is a Krylov space?
	  \end{Problem}

	  \begin{Problem}
		  What is the difference between the Arnoldi and the Lanczos processes?
	  \end{Problem}

	  \begin{Problem}
		  Is there a short recurrence for the GMRES method?
	  \end{Problem}

	  \begin{Problem}
		  What is preconditioning?
	  \end{Problem}

	  \begin{Problem}
		  What complicates preconditioning for the conjugate gradient method?
	  \end{Problem}

\end{Sheet}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End: 
